\chapter*{List of related publications}
\addcontentsline{toc}{chapter}{List of related publications}

\subsection*{Articles}
\textbf{Duy-Nam Bui}, Manh Duong Phung, and Hung Pham Duy. ``Event-based Reconfiguration Control for Time-varying Formation of Robot Swarms in Narrow Spaces,'' \textit{Manuscript}.

\textbf{Duy-Nam Bui}, Manh Duong Phung, and Hung Pham Duy. ``Predictive Reconfiguration Control for Multi-Robot Formation in Cluttered Environments,'' \textit{Manuscript}.

\textbf{Duy-Nam Bui} and Manh Duong Phung, ``Radial Basis Function Neural Networks for Formation Control of Unmanned Aerial Vehicles,'' in \textit{Robotica}, vol. 42, pp. {1842--1860}, 2024.

Thi Thuy Ngan Duong, \textbf{Duy-Nam Bui}, and Manh Duong Phung, ``Navigation Variable-based Multi-objective Particle Swarm Optimization for UAV Path Planning with Kinematic Constraints,'' in \textit{Neural Computing and Applications}, 2024.

\textbf{Duy-Nam Bui}, Manh Duong Phung and Hung Pham Duy, ``Self-Reconfigurable V-Shape Formation of Multiple UAVs in Narrow Space Environments,'' \textit{2024 IEEE/SICE International Symposium on System Integration (SII)}, Ha Long, Vietnam, pp. 1006--1011, 2024.

\textbf{Duy-Nam Bui}, Thuy Ngan Duong and Manh Duong Phung, ``Ant Colony Optimization for Cooperative Inspection Path Planning Using Multiple Unmanned Aerial Vehicles,'' \textit{2024 IEEE/SICE International Symposium on System Integration (SII)}, Ha Long, Vietnam, pp. 675--680, 2024.

\textbf{Duy-Nam Bui}, Thu Hang Khuat, Manh Duong Phung, Thuan-Hoang Tran, Dong LT Tran, ``Optimal Motion Planning for Unmanned Aerial Vehicles in Unknown Environments,'' \textit{2024 International Conference on Control, Robotics and Informatics (ICCRI)}, Da Nang, Vietnam, 2024.
% Thu Hang Khuat, \textbf{Duy-Nam Bui}, Thuy Ngan Duong, and Manh Duong Phung, ``Polar Coordinate-based Differential Evolution for Moving Target Search Using Vision Sensor on Unmanned Aerial Vehicles,'' \textit{Robotics and Autonomous Systems}.
%     Thu Hang Khuat, \textbf{Duy-Nam Bui}, Hoa TT. Nguyen, Mien L. Trinh, Minh T. Nguyen, and Manh Duong Phung, ``Multi-goal Rapidly Exploring Random Tree with Safety and Dynamic Constraints for UAV Cooperative Path Planning,'' \textit{IEEE Transactions on Vehicular Technology}.


\subsection*{Open-source software}
UAVs simulator -- {\tt\url{https://github.com/duynamrcv/hummingbird_simulator}}.

NMPC controller -- {\tt\url{https://github.com/duynamrcv/hummingbird_nmpc}}.

SRVC strategy -- {\tt\url{https://github.com/duynamrcv/reconfigurable_vshape}}.

ERC strategy -- {\tt\url{https://github.com/duynamrcv/erc}}.

PRC strategy -- {\tt\url{https://github.com/duynamrcv/prc}}.