\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
% \fontsize{13}{15}\selectfont

This master's thesis would not have been possible without the help, support, and contributions from numerous people. To begin with, I would like to express my sincere gratitude to Dr. Pham Duy Hung, lecturer of the Automatic control and Robotics Lab at VNU University of Engineering and Technology, who gave me a rare opportunity to conduct my master's studies within the lab. Thank you for your enthusiastic guidance during my program, your trust in my abilities, and for inspiring me. His patient yet strict guidance helped me broaden my horizons significantly. His enthusiasm, endless passion, and professional attitude towards science and research also emerged as a considerable source of motivation for me to overcome all difficulties in research and academic activities. I am also really grateful to Dr. Phung Manh Duong for giving me a hand in finding my true passion for my academic career. His contributions, advice, and encouragement helped me to enhance my knowledge in leaps and bounds and refine my ideas during the research process and completion of my thesis. His enthusiastic guidance helped me firmly step forward in the first steps of my research career.

In addition, I would also like to thank members of the Automatic control and Robotics Lab for the time we spent together, and for the invaluable inspiration and help with numerous experiments and publications. I also want to express my appreciation to all members, colleagues, teachers, and friends from the Faculty of Electronics and Telecommunication for all of your love and support. Finally, I would like to thank all the people who reviewed this thesis and for their honest and useful feedback.

And most importantly, I would like to express my deepest gratitude to my beloved family and friends, who were always there when I needed them, especially my mother, and my father, for their unconditional love, and for always having supported me in pursuing my plans, and forever the most peaceful place for me to lean on.

\vspace{0.5cm}
\begin{flushleft}
Bui Duy Nam
\end{flushleft}

\subsection*{Financial Support}
The research leading to the publications and results presented in this thesis was supported by the Master, PhD Scholarship Programme of the Vingroup Innovation Foundation (VINIF), codes VINIF.2022.Ths.057 and VINIF.2023.Ths.088.